---
title: "R Notebook"
author: "Elena Lazarenko"
output:
  html_document:
    df_print: paged
---

This is my mini-project for finance course where I decided to do some portfolio optimization. 

```{r}
library(timeSeries)
library(fPortfolio)
library(quantmod)
library(caTools)
library(dplyr)
library(PerformanceAnalytics)
library(ggplot2)
library(corrplot)
```

I took absolutely different stocks: Apple and Amazon which I personally like, Cray Inc and CAE Inc which have up on unusual volume, Baidu, Inc. and Coca-Cola Consolidated, Inc. which stock prices are going down.
```{r}
# Creating Vector of Tickers
tickers <- c("AAPL", "NFLX", "CRAY", "CAE", "BIDU", "COKE")
```

Now I am trying to get prices of stocks since January 1, 2017
```{r}
# Get daily returns
prices <- NULL
for (t in tickers)
  prices <- cbind(prices, getSymbols(t, from="2017-01-01", src ="yahoo", auto.assign=FALSE)[,4])
```

Oh, some values are missing! We can delete the dates where we don't have prices
```{r}
prices <- prices[apply(prices, 1, function(x) all(!is.na(x))),]
colnames(prices) <- tickers
```

```{r}
# Calculating daily Return on Capital
preturns <- na.omit(ROC(prices, type="discrete"))
preturns <- as.timeSeries(prices)
```

```{r}
# Calculate Monthly Returns
sdata <- tickers %>% lapply(function(x) getSymbols(x, from="2017-01-01", src="yahoo", auto.assign=FALSE)[,4]) %>% lapply(function(x) monthlyReturn(x))
```


```{r}
preturns <- do.call(merge, sdata)
# delete dates that do not have closing prices for all tickers
preturns <- preturns[apply(preturns, 1, function(x) all(!is.na(x))),]
colnames(preturns) <- tickers
preturns <- as.timeSeries(preturns)
```

Now I will calculate th efficient frontier. The efficient frontier is the set of optimal portfolios that offer the highest expected return for a defined level of risk or the lowest risk for a given level of expected return. Portfolios that lie below the efficient frontier are sub-optimal because they do not provide enough return for the level of risk. Portfolios that cluster to the right of the efficient frontier are sub-optimal because they have a higher level of risk for the defined rate of return.
```{r}
# calculate the efficient frontier
efrontier <- portfolioFrontier(preturns, constraints = "LongOnly")
efrontier
```

Let's plot efficient frontier for better understanding.
```{r}
# plot frontier
plot(efrontier,c(1,2,3,4))
```

Let's see whether stock prices of different companies are correlated.
```{r}
# Plot correlation matrix
cormatrix <- cor(preturns)
corrplot(cormatrix, method="circle")
```

Now let's build the covariance matrix. If some stock prices increase and decrease simultaneously then the covariance value will be positive. Conversely if some increase while the other decrease then the covariance will be negative. As you can see covariancw is almost zero everywhere, I even can not visualize it.
```{r}
# Covariance matrix
covmatrix <- cov(preturns)
covmatrix
```

```{r}
#Turn data to annual
rrpoints <- frontierPoints(efrontier)
anpoints <- data.frame(targetRisk=rrpoints[, "targetRisk"] * sqrt(252),
                               targetReturn=rrpoints[,"targetReturn"] * 252)
plot(anpoints)
```

```{r}
# Sharpe ratios for each point on the efficient frontier
risk_free_rate <- 0
plot((anpoints[,"targetReturn"]-risk_free_rate) / anpoints[,"targetRisk"], xlab="point on efficient frontier", ylab="Sharpe ratio")
```

Here I defined Minimum Variance Portfolio. MVP means a diversified portfolio that mitigates risk properly.
```{r}
# Minimum Variance Portfolio
mvp <- minvariancePortfolio(preturns, spec=portfolioSpec(), constraints="LongOnly")
mvp
```

And that is the Tangency Portfolio. Tangency portfolio is the result of the intercept point of Capital Market Line and efficient frontier
```{r}
# Tangency Portfolio
tp <- tangencyPortfolio(preturns, spec=portfolioSpec(), constraints="LongOnly")
tp
```

Let's see what are the weights of both portfolios and visualize them.

```{r}
mvpweights <- getWeights(mvp)
tangencyweights <- getWeights(tp)
```


```{r}
 
# Minimun Variance Portfolio Weights
df <- data.frame(mvpweights)
assets <- colnames(fweights)
ggplot(data=df, aes(x=assets, y=mvpweights, fill=assets)) +
  geom_bar(stat="identity", position=position_dodge(),colour="black") +
  geom_text(aes(label=sprintf("%.02f %%",mvpweights*100)),
            position=position_dodge(width=0.9), vjust=-0.25, check_overlap = TRUE) +
              ggtitle("Minimum Variance Portfolio Optimal Weights")+ theme(plot.title = element_text(hjust = 0.5)) +
                labs(x= "Assets", y = "Weight (%)")
```

```{r}
# Tangency Portfolio Weights
dft <- data.frame(tangencyweights)
assets <- colnames(fweights)
ggplot(data=dft, aes(x=assets, y=tangencyweights, fill=assets)) +
  geom_bar(stat="identity", position=position_dodge(),colour="black") +
  geom_text(aes(label=sprintf("%.02f %%",tangencyweights*100)),
            position=position_dodge(width=0.9), vjust=-0.25, check_overlap = TRUE) +
  ggtitle("Tangency Portfolio Weights")+ theme(plot.title = element_text(hjust = 0.5)) +
  labs(x= "Assets", y = "Weight (%)")
```


```{r}
# Pie of Minimum Variance Portfolio Weight
bar <- ggplot(df, aes(x = "", y = mvpweights, fill=assets)) + geom_bar(width= 1, stat="identity") + ggtitle("Minimum Variance Portfolio Weights")+ theme(plot.title = element_text(hjust = 0.5)) 
pie <- bar + coord_polar("y", start=0)
pie + scale_fill_brewer(palette="Blues")+
  theme_minimal()
```

```{r}
# Pie of Tangency Portfolio Weights
bar <- ggplot(dft, aes(x = "", y = tangencyweights, fill=assets)) + geom_bar(width= 1, stat="identity") + ggtitle("Tangency Portfolio Weights")+ theme(plot.title = element_text(hjust = 0.5)) 
pie <- bar + coord_polar("y", start=0)
pie + scale_fill_brewer(palette="Blues")+
  theme_minimal()
```

Actually after all of it I can not tell which portfolio I would personally choose. Maybe I would rather prefer Tangency Portfolio because I think that this is less risky than the Minimum Variance Portfolio.I see that Minimum Variance portfolio suggest to invest in Baidu but now their stock prices are going down. And both Tangency Portfolio and MVP suggest to invest in CAE Inc the biggest weight which is interesting because now their stock prices are going up very fast. So from two plots I can tell that I would rather invest in Netflix than in Baidu although MVP shows vice versa but I don't think that it is right way to act.
