---
title: "R Notebook"
author: "Elena Lazarenko"
output:
  html_document:
    df_print: paged
---

This is my mini-project for finance course where I decided to do some portfolio optimization. 

```{r}
library(timeSeries)
library(fPortfolio)
library(quantmod)
library(caTools)
library(dplyr)
library(PerformanceAnalytics)
library(ggplot2)
library(corrplot)
```

I took absolutely different stocks: Apple and Amazon which I personally like, Cray Inc and CAE Inc which have up on unusual volume, Baidu, Inc. and Coca-Cola Consolidated, Inc. which stock prices are going down.
```{r}
# Creating Vector of Tickers
tickers <- c("AAPL", "NFLX", "CRAY", "CAE", "BIDU", "COKE")
```

Now I am trying to get prices of stocks since January 1, 2017
```{r}
# Get daily returns
prices <- NULL
for (t in tickers)
  prices <- cbind(prices, getSymbols(t, from="2017-01-01", src ="yahoo", auto.assign=FALSE)[,4])
```

Oh, some values are missing! We can delete the dates where we don't have prices
```{r}
prices <- prices[apply(prices, 1, function(x) all(!is.na(x))),]
colnames(prices) <- tickers
```

```{r}
# Calculating daily Return on Capital
preturns <- na.omit(ROC(prices, type="discrete"))
preturns <- as.timeSeries(prices)
```

```{r}
# Calculate Monthly Returns
sdata <- tickers %>% lapply(function(x) getSymbols(x, from="2017-01-01", src="yahoo", auto.assign=FALSE)[,4]) %>% lapply(function(x) monthlyReturn(x))
```

```{r}
preturns <- do.call(merge, sdata)
# delete dates that do not have closing prices for all tickers
preturns <- preturns[apply(preturns, 1, function(x) all(!is.na(x))),]
colnames(preturns) <- tickers
preturns <- as.timeSeries(preturns)
```

```{r}
# calculate the efficient frontier
efrontier <- portfolioFrontier(preturns, constraints = "LongOnly")
efrontier
```

```{r}
# plot frontier
plot(efrontier,c(1,2,3,4))
```


```{r} 
#Plot Frontier Weights
fweights <- getWeights(efrontier)
colnames(fweights) <- tickers
risk_return <- frontierPoints(efrontier)
```

Let's see whether stock prices of different companies are correlated.
```{r}
# Plot correlation matrix
cormatrix <- cor(preturns)
corrplot(cormatrix, method="circle")
```

```{r}
# Covariance matrix
covmatrix <- cov(preturns)
covmatrix
```

```{r}
#Turn data to annual
rrpoints <- frontierPoints(efrontier)
anpoints <- data.frame(targetRisk=rrpoints[, "targetRisk"] * sqrt(252),
                               targetReturn=rrpoints[,"targetReturn"] * 252)
plot(anpoints)
```

```{r}
# Sharpe ratios for each point on the efficient frontier
risk_free_rate <- 0
plot((anpoints[,"targetReturn"]-risk_free_rate) / anpoints[,"targetRisk"], xlab="point on efficient frontier", ylab="Sharpe ratio")
```

```{r}
# Frontier Weights
barplot(t(fweights), main="Frontier Weights", col=cm.colors(ncol(fweights)+2), legend=colnames(fweights))
```


```{r}
#Minimum Variance Portfolio
mvp <- minvariancePortfolio(preturns, spec=portfolioSpec(), constraints="LongOnly")
mvp
```

```{r}
# Tangency Portfolio
tp <- tangencyPortfolio(preturns, spec=portfolioSpec(), constraints="LongOnly")
tp
```


```{r}
mvpweights <- getWeights(mvp)
tangencyweights <- getWeights(tp)
```


```{r}
 
#Extract value at risk
covRisk(preturns, mvpweights)
varRisk(preturns, mvpweights, alpha = 0.05)
cvarRisk(preturns, mvpweights, alpha = 0.05)
```

```{r}
 
# Minimun Variance Portfolio Weights
df <- data.frame(mvpweights)
assets <- colnames(fweights)
ggplot(data=df, aes(x=assets, y=mvpweights, fill=assets)) +
  geom_bar(stat="identity", position=position_dodge(),colour="black") +
  geom_text(aes(label=sprintf("%.02f %%",mvpweights*100)),
            position=position_dodge(width=0.9), vjust=-0.25, check_overlap = TRUE) +
              ggtitle("Minimum Variance Portfolio Optimal Weights")+ theme(plot.title = element_text(hjust = 0.5)) +
                labs(x= "Assets", y = "Weight (%)")
```

```{r}
# Tangency Portfolio Weights
dft <- data.frame(tangencyweights)
assets <- colnames(fweights)
ggplot(data=dft, aes(x=assets, y=tangencyweights, fill=assets)) +
  geom_bar(stat="identity", position=position_dodge(),colour="black") +
  geom_text(aes(label=sprintf("%.02f %%",tangencyweights*100)),
            position=position_dodge(width=0.9), vjust=-0.25, check_overlap = TRUE) +
  ggtitle("Tangency Portfolio Weights")+ theme(plot.title = element_text(hjust = 0.5)) +
  labs(x= "Assets", y = "Weight (%)")
```


```{r}
# Pie of Minimum Variance Portfolio Weight
bar <- ggplot(df, aes(x = "", y = mvpweights, fill=assets)) + geom_bar(width= 1, stat="identity") + ggtitle("Minimum Variance Portfolio Weights")+ theme(plot.title = element_text(hjust = 0.5)) 
pie <- bar + coord_polar("y", start=0)
pie + scale_fill_brewer(palette="Blues")+
  theme_minimal()
```

```{r}
# Pie of Tangency Portfolio Weights
bar <- ggplot(dft, aes(x = "", y = tangencyweights, fill=assets)) + geom_bar(width= 1, stat="identity") + ggtitle("Tangency Portfolio Weights")+ theme(plot.title = element_text(hjust = 0.5)) 
pie <- bar + coord_polar("y", start=0)
pie + scale_fill_brewer(palette="Blues")+
  theme_minimal()
```



